{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion}

### 2017-01-05 proposal/overview.tex
minor fixes

{\bf Preemption points.}
During execution, a model checker identifies a subset of the program's operations as ``interesting'', i.e.,
points at which interrupting the current thread to run a different one is likely to produce different behaviour.
These points, called {\em preemption points}, may be identified by any combination of human intuition and machine analysis.
Typical preemption points include the boundaries of synchronization APIs (e.g., {\tt mutex\_lock}) or accesses to shared variables.
Considering that at each preemption point multiple threads exist as options to run next, the set of possible ways to execute the program can be viewed as a tree.
% TODO: fig tree
I show an example execution tree in Figure~\ref{fig:tree}.
The number of preemption points in each execution define the depth of this tree, and the number of threads available to run define the branching factor.
Hence, in a program with $n$ preemption points and $k$ threads available to run at each, the state space size is $n^k$.
Addressing the scaling problem in this exponential relation is the central research problem for all model checkers.

@@

A data race analysis may be either {\em static} (inspecting source code) \cite{racerx} or {\em dynamic} (tracking individual accesses arising at run-time) \cite{tsan}.
This paper focuses exclusively on dynamic analysis,
so although our example refers to numbered source lines for ease of explanation,
in practice we are actually classifying the individual memory access events corresponding to those lines during execution.
Actually, each {\tt x++} statement likely compiles to two separate load and store instructions, so each of those two instructions from each of the two marked source lines pairwise will race (except for the two loads, which are both reads).

>>>>

{\bf Preemption points.}
During execution, a model checker identifies a subset of the program's operations as ``interesting'', i.e.,
points at which interrupting the current thread to run a different one is likely to produce different behaviour.
These points, called {\em preemption points}, may be identified by any combination of human intuition and machine analysis.
Typical preemption points include the boundaries of synchronization APIs (e.g., {\tt mutex\_lock}) or accesses to shared variables.
Considering that at each preemption point multiple threads exist as options to run next, the set of possible ways to execute the program can be viewed as a tree.
% TODO: fig tree
I show an example execution tree in Figure~\ref{fig:tree}.
The number of preemption points in each execution define the depth of this tree, and the number of threads available to run define the branching factor.
Hence, in a program with $n$ preemption points and $k$ threads available to run at each, the state space size is $O(n^k)$.
Addressing the scaling problem in this exponential relation is the central research problem for all model checkers.

@@

A data race analysis may be either {\em static} (inspecting source code) \cite{racerx} or {\em dynamic} (tracking individual accesses arising at run-time) \cite{tsan}.
This paper focuses exclusively on dynamic analysis,
so although our example refers to numbered source lines for ease of explanation,
in practice we are actually classifying the individual memory access events corresponding to those lines during execution.
Actually, each {\tt x++} statement likely compiles to two separate load or store instructions, so each of those two instructions from each of the two marked source lines pairwise will race (except for the two loads, which are both reads).
